Additional analysis

### 2018-07-14 03_additional-analysis.Rmd
Use Bayesian inference for sample vs. population comparisons
THIS IS SO COOL.

# Load libraries again
library(tidyverse)
library(magrittr)
library(rstanarm)
library(broom)
library(coda)
library(glue)
library(ggstance)
library(ggridges)
library(grid)
library(gridExtra)
library(Gmisc)
library(patchwork)
library(pander)
library(scales)
library(huxtable)
library(lubridate)
library(modelr)
library(here)

@@

@@

coef_summary_charity <- newdata_conditions_charity %>% 
  bind_cols(tidyMCMC(as.mcmc(chains_fitted_charity), estimate.method = "median",
                     conf.int = TRUE, conf.level = 0.9, conf.method = "HPDinterval")) %>% 
  mutate_at(vars(crackdown, give_charity_3), funs(fct_inorder(., ordered = TRUE)))

@@

coef_summary_vol_relig <- newdata_conditions_vol_relig %>% 
  bind_cols(tidyMCMC(as.mcmc(chains_fitted_vol_relig), estimate.method = "median",
                     conf.int = TRUE, conf.level = 0.9, conf.method = "HPDinterval")) %>% 
  mutate_at(vars(crackdown, volunteer, religiosity_bin), funs(fct_inorder(., ordered = TRUE))) %>% 
  mutate(religiosity_bin = fct_recode(religiosity_bin,
                                      `Rarely attend religious services` = "Rarely",
                                      `Attend religious services\nat least once a month` = 
                                        "At least once a month"))

@@

m_likely_cif_full <-update(m_likely_cif, . ~ . + 
                             favor_humanitarian_bin +
                             give_charity_3 + volunteer + political_knowledge_bin +
                             ideology_bin + education_bin + religiosity_bin +
                             income_bin + age_bin)
```

>>>>

# Load libraries again
library(tidyverse)
library(magrittr)
library(rstan)
library(rstanarm)
library(broom)
library(glue)
library(ggstance)
library(ggridges)
library(grid)
library(gridExtra)
library(Gmisc)
library(patchwork)
library(pander)
library(scales)
library(huxtable)
library(lubridate)
library(modelr)
library(ipumsr)
library(here)

@@

# Characteristics of experiment samples

We compare our sample with demographic characteristics of the general population. Since there is no nationally representative sample for each of our demographic variables, we look at multiple data sources.

- Age:
- Volunteering
- Donating to charity:

Extract data from Current Population Survey, housed at the Minnesota Population Center’s Integrated Public Use Microdata Series

```{r load-clean-cps, warning=FALSE, message=FALSE}
cps_2015_ddi_file <- file.path(here(), "data", "ipums-cps", "cps_09_2015.xml")
cps_2015_data_file <- file.path(here(), "data", "ipums-cps", "cps_09_2015.dat.gz")

cps_2015_ddi <- read_ipums_ddi(cps_2015_ddi_file)
cps_2015_data <- read_ipums_micro(cps_2015_ddi_file, data_file = cps_2015_data_file, verbose = FALSE)

cps_2017_ddi_file <- file.path(here(), "data", "ipums-cps", "cps_2017.xml")
cps_2017_data_file <- file.path(here(), "data", "ipums-cps", "cps_2017.dat.gz")

cps_2017_ddi <- read_ipums_ddi(cps_2017_ddi_file)
cps_2017_data <- read_ipums_micro(cps_2017_ddi_file, data_file = cps_2017_data_file, verbose = FALSE)

# Volunteering data from September 2015 only
df_volunteering <- cps_2015_data %>% 
  # Remove values not in the universe
  mutate_at(vars(VLSTATUS, VLDONATE), funs(ifelse(. == 99, NA, .)))

# All other data from annual March 2017 survey
df_demographics <- cps_2017_data %>% 
  # Remove values not in the universe
  mutate(SEX = ifelse(SEX == 9, NA, SEX),
         EDUC = ifelse(EDUC <= 1 | EDUC == 999, NA, EDUC),
         INCTOT = ifelse(INCTOT == 99999999, NA, INCTOT))
```

```{r population-values}
global_demographics <- df_demographics %>% 
  summarize(age = weighted.mean(AGE >= 35, ASECWT), 
            female = weighted.mean(SEX == 2, ASECWT),
            college = weighted.mean(EDUC >= 111, ASECWT, na.rm = TRUE),
            income = weighted.mean(INCTOT >= 50000, ASECWT, na.rm = TRUE)) %>% 
  c()

global_vol <- df_volunteering %>% 
  summarize(volunteering = weighted.mean(VLSTATUS == 2, VLSUPPWT, na.rm = TRUE),
            donating = weighted.mean(VLDONATE == 2, VLSUPPWT, na.rm = TRUE)) %>% 
  c()

global_stats <- c(global_vol, global_demographics)
```

```{r sample-population-characteristics-freq, include=FALSE, eval=FALSE}
# Demographic variables
table(results$gender_bin) %>% prop.test(., p = global_stats$female)
table(results$education_bin) %>% prop.test(., p = global_stats$college)
table(results$income_bin) %>% prop.test(., p = global_stats$income)
table(results$age_bin) %>% prop.test(., p = global_stats$age)

# Volunteering
table(results$volunteer) %>% prop.test(., p = global_stats$volunteering)
table(results$give_charity_2) %>% prop.test(., p = global_stats$donating)
```

```{r compile-stan-model, warning=FALSE, message=FALSE, cache=TRUE}
prop_test_bayes_saved <- file.path(here(), "lib", "bayes_prop_test.rds")

if (file.exists(prop_test_bayes_saved)) {
  prop_test_bayes <- readRDS(prop_test_bayes_saved)
} else {
  prop_test_bayes <- stan_model(file.path(here(), "lib", "bayes_prop_test.stan"))
  saveRDS(prop_test_bayes, prop_test_bayes_saved)
}
```

```{r sample-population-characteristics, cache=TRUE}
compare_sample_to_pop_bayes <- function(sample_value, population_value) {
  mcmc_samples <- sampling(prop_test_bayes, list(x = table(sample_value)[1],
                                                 total_n = length(sample_value),
                                                 pop_prop = population_value),
                           chains = CHAINS, iter = ITER, warmup = WARMUP, seed = BAYES_SEED)

  tidied <- tidyMCMC(mcmc_samples, conf.int = TRUE, conf.level = 0.9, 
                     estimate.method = "median", conf.method = "HPDinterval") %>%
    mutate(in_hpdi = (population_value >= conf.low & population_value <= conf.high))

  thetas <- unlist(extract(mcmc_samples, "theta"))
  pop_quantile_in_sample <- ecdf(thetas)(population_value)

  in_hpdi <- (population_value >= tidied[1,]$conf.low & 
                population_value <= tidied[1,]$conf.high)

  return(list(mcmc_samples = mcmc_samples, tidied = tidied, theta_in_hpdi = in_hpdi,
              pop_quantile_in_sample = pop_quantile_in_sample))
}

calc_sample_pop <- tribble(
  ~Variable, ~sample_value, ~National,
  "Female (%)", results$gender_bin, global_stats$female,
  "Age (% 35+)", results$age_bin, global_stats$age,
  "Income (% $50,000+)", results$income_bin, global_stats$income,
  "Education (% BA+)", results$education_bin, global_stats$college,
  "Donated in past year (%)", results$give_charity_2, global_stats$donating,
  "Volunteered in past year (%)", results$volunteer, global_stats$volunteering
) %>% 
  mutate(Sample = sample_value %>% map_dbl(~ prop.table(table(.))[1]),
         prop_test_bayes = map2(.x = sample_value, .y = National, 
                                .f = ~ compare_sample_to_pop_bayes(.x, .y))) 
```

```{r tbl-sample-characteristics, results="asis"}
format_hpdi <- function(post_lower, post_upper, star, digits = 1) {
  glue("({lower}%, {upper}%){star}",
       lower = round(100 * post_lower, digits),
       upper = round(100 * post_upper, digits))
}

tbl_sample_pop <- calc_sample_pop %>% 
  mutate(in_hpdi = prop_test_bayes %>% map_lgl(~ .$theta_in_hpdi),
         not_hpdi_symbol = ifelse(in_hpdi, "", "†"),
         diffs_tidy = prop_test_bayes %>% map(~ .$tidied[2,]),
         diffs_median = diffs_tidy %>% map_dbl(~ .$estimate),
         diffs_hpdi_fancy = diffs_tidy %>%
           map2_chr(.x = diffs_tidy, .y = not_hpdi_symbol, 
                    .f = ~ format_hpdi(.x$conf.low, .x$conf.high, .y))) %>% 
  mutate_at(vars(National, Sample, diffs_median), funs(percent)) %>% 
  select(Variable, Sample, National, 
         `Sample − National~posterior\\ median~` = diffs_median,
         `90% HPDI` = diffs_hpdi_fancy)

tbl_sample_pop %>% 
  pandoc.table.return(keep.line.breaks = TRUE, style = "multiline", justify = "lcccc", 
                      caption = "Characteristics of experimental sample {#tbl:exp-sample}") %T>% 
  cat(file = file.path(here(), "output", "tables", "tbl-exp-sample.md")) %>% 
  cat()
```

@@

coef_summary_charity <- newdata_conditions_charity %>% 
  bind_cols(tidyMCMC(chains_fitted_charity, estimate.method = "median",
                     conf.int = TRUE, conf.level = 0.9, conf.method = "HPDinterval")) %>% 
  mutate_at(vars(crackdown, give_charity_3), funs(fct_inorder(., ordered = TRUE)))

@@

coef_summary_vol_relig <- newdata_conditions_vol_relig %>% 
  bind_cols(tidyMCMC(chains_fitted_vol_relig, estimate.method = "median",
                     conf.int = TRUE, conf.level = 0.9, conf.method = "HPDinterval")) %>% 
  mutate_at(vars(crackdown, volunteer, religiosity_bin), funs(fct_inorder(., ordered = TRUE))) %>% 
  mutate(religiosity_bin = fct_recode(religiosity_bin,
                                      `Rarely attend religious services` = "Rarely",
                                      `Attend religious services\nat least once a month` = 
                                        "At least once a month"))

@@

m_likely_cif_full <- update(m_likely_cif, . ~ . + 
                              favor_humanitarian_bin +
                              give_charity_3 + volunteer + political_knowledge_bin +
                              ideology_bin + education_bin + religiosity_bin +
                              income_bin + age_bin)
```
